---
title: "Material and methods"
author: "Rémi Gau"
date: "2022-06-05"
output: word_document
# output: html_document
bibliography: references.bib
---


------

# Material and methods

<!--
Ethics approval
-->


## Sample description

TODO add participants descritption table

<!--
```{r, echo=FALSE, warning=FALSE}
library(flextable) #you can use another table library, if you want
flextable(read.csv("participants.csv"))
```
-->

## Behavioral testing (out of scan)

## Experimental design

## Stimuli description

## MRI acquisition parameters

```{r, child=c('sub-ctrl01/bids-matlab_report.md')}
```
## Quality control

```{r, child=c('quality_control.md')}
```

## Preprocessing

```{r, child=c('fmriprep/CITATION.md')}
```

## Statistical analysis

### Behavioral

### fMRI


```{r, child=c('statistical_analysis_fmri.md')}
```

## Data and code

Code and data are bundled together as a datalad dataset available on GIN:
https://gin.g-node.org/cpp-lln-lab/olf_blind_yoda

Code is also available on [Github](https://github.com/Remi-Gau/olfactory_fmri_blind).

Raw behavioral in-scan data are available on Gin : https://gin.g-node.org/cpp-lln-lab/olfaction_blind_beh

Raw and derivative (fMRIprep and MRIQC) output data is available on requests.

<!--
where to store stats data at the group level
-->

Statiscal maps are availble on neurovault

## References
